Last glacial cycle glacier erosion potential in the Alps

### 2021-02-04 articles/erosion/alps-erosion.tex
Rework sensitivity sections with Ian's comments.

    The results found on the Rhine Glacier catchment can be generalized to the
    entire model domain by using (present-day) bedrock altitude as a proxy for
    along-flow distance (Fig.~\ref{fig:hypsogram}). Thus more generally,
    periods of modelled increasing and maximum ice volume corresponds to lower
    values for elevation-aggregated erosion rates, with significant erosion
    restricted to lower elevations. On the opposite, periods of modelled
    decreased ice volume correspond to higher local modelled erosion rates and
    more extensive rapid erosion into higher-elevation areas
    (Fig.~\ref{fig:hypsogram}).

@@

    Different paleoclimate forcing scenarios lead to different modelled
    glaciation histories and thus different results in terms of erosion
    potential (Fig.~\ref{fig:sensitivity}). Scenarios including higher
    (constant) input precipitation rates yield generally higher modelled
    erosion potential (Fig.~\ref{fig:sensitivity}a, b, and~c), while scenarios
    including paleo-precipitation reductions (Fig.~\ref{fig:sensitivity}d, e,
    and~f), including our default high-resolution run, yield a smaller erosion
    potential. The spatial pattern of modelled erosion potential is
    nevertheless generally consistent across paleoclimate scenarios tested,
    despite variations in the glaciation maximum extent
    (Fig.~\ref{fig:sensitivity}). Furthermore, these changes are minor
    considering the large uncertainties affecting the choice of erosion law.

@@

    The choice of erosion law significantly impacts the results
    (Fig.~\ref{fig:powerlaws}). Our default erosion law, based on quantified
    sediment yields from Patagonian and Antarctic Peninsula tidewater glaciers
    \citep[${\dot{e}=5.2\times10^{-8}u_\mathrm{b}^{2.34}}$,][]{Koppes.etal.2015}
    yields a moderate and strongly localized cumulative erosion potential
    (Figs~\ref{fig:cumulative} and~\ref{fig:powerlaws}a). The erosion law
    based on measured suspended sediment load from Franz-Joseph Glacier
    \citep[${\dot{e}=2.7\times10^{-7}u_\mathrm{b}^{2.02}}$,][]{Herman.etal.2015}
    also yields a strongly localized, and yet much higher, modelled erosion
    potential (Fig.~\ref{fig:powerlaws}b). The erosion law based on measured
    suspended sediment load during a surge of Variegated Glacier
    \citep[${\dot{e}=1\times10^{-4}u_\mathrm{b}^{1}}$,][]{Humphrey.Raymond.1994},
    results in similarly high values, but a less localized pattern of cumulative
    erosion potential (Fig.~\ref{fig:powerlaws}c). Finally, The erosion law
    based on a global compilation of glacier velocity and erosion rates
    \citep[${\dot{e}=1.665\times10^{-1}u_\mathrm{b}^{0.6459}}$,][]{Cook.etal.2020}
    yields an even flatter pattern of cumulative erosion potential
    (Fig.~\ref{fig:powerlaws}d).

    It may seem surprising that the erosion law calibrated on Patagonian and
    Antarctic Peninsula tidewater glaciers \citep{Koppes.etal.2015} yields a
    more realistic cumulative erosion potential than those based on terrestrial
    glaciers \citep{Humphrey.Raymond.1994, Herman.etal.2015, Cook.etal.2020},
    themselves resulting
    in cumulative erosion potential values in the order of kilometres and thus
    higher than the Alpine relief itself. However, during the Last Glacial
    Maximum and much of the last glacial cycle, Alpine paleoglaciers were
    closer in size and climatic context to the present-day glaciers of
    Patagonia and the Antarctic Peninsula \citep{Koppes.etal.2015} than to
    Franz-Joseph Glacier \citep{Herman.etal.2015} and many of the glaciers
    included in the global compilation by \citet{Cook.etal.2020}. Atmospheric
    temperatures have been found to be a strong driver of erosion in studies
    involving several glaciers \citep{Koppes.etal.2015, Cook.etal.2020}.

    That said, the magnitude of the modelled erosion rates can hardly be used
    as a criteria to rule out this or that erosion law for the glacial Alps, as
    all these laws are affected with very large uncertainties. It can not be
    excluded that a linear or sub-linear erosion-law
    \citep[e.g.][]{Cook.etal.2020} would lead to a less localized pattern of
    erosion (Fig.~\ref{fig:powerlaws}d), and a stronger correlation between
    modelled ice volume and erosion (Fig.~\ref{fig:powerlaws}l). Even
    then, the intrinsically large variations in glacier sliding velocities
    yields modelled glacial-cycle erosion potential two to three orders of
    magnitude higher in the Alpine valley than near the mountain tops
    (Fig.~\ref{fig:powerlaws}d).

>>>>

    The results found on the Rhine Glacier catchment can be generalized to the
    entire model domain by using (present-day) bedrock altitude as a proxy for
    along-flow distance (Fig.~\ref{fig:hypsogram}). Thus more generally,
    periods of modelled increasing and maximum ice volume correspond to lower
    values for elevation-aggregated erosion rates, with significant erosion
    restricted to lower elevations. On the opposite, periods of modelled
    decreased ice volume correspond to higher local modelled erosion rates and
    more extensive rapid erosion into higher-elevation areas
    (Fig.~\ref{fig:hypsogram}).

@@

    Different paleoclimate forcing scenarios lead to different modelled
    glaciation histories \citep[cf.][]{Seguinot.etal.2018} and thus different
    results in terms of erosion
    potential (Fig.~\ref{fig:sensitivity}). Scenarios including higher
    (constant, present-day) input precipitation rates yield generally higher
    modelled
    erosion potential (Fig.~\ref{fig:sensitivity}a, b, and~c), while scenarios
    including paleo-precipitation reductions (Fig.~\ref{fig:sensitivity}d, e,
    and~f), including our default high-resolution run
    (Fig.~\ref{fig:cumulative}), yield a smaller erosion potential. The spatial
    pattern of modelled time-integrated erosion is nevertheless generally
    consistent across the paleoclimate scenarios tested,
    despite variations in the glaciation maximum extent
    (Fig.~\ref{fig:sensitivity} and \citealp[Fig.~3 in][]{Seguinot.etal.2018}).
    Furthermore, these changes are minor
    considering the large uncertainties affecting the choice of erosion law.

@@

    The choice of erosion law significantly impacts the results
    (Fig.~\ref{fig:powerlaws}). Our default, based on quantified
    sediment yields from Patagonian and Antarctic Peninsula tidewater glaciers
    \citep[${\dot{e}=5.2\times10^{-8}u_\mathrm{b}^{2.34}}$,][]{Koppes.etal.2015}
    yields a moderate and strongly localized cumulative erosion potential
    (Figs~\ref{fig:cumulative} and~\ref{fig:powerlaws}a). The erosion law
    based on measured suspended sediment load from Franz-Joseph Glacier
    \citep[${\dot{e}=2.7\times10^{-7}u_\mathrm{b}^{2.02}}$,][]{Herman.etal.2015}
    also yields a strongly localized, and yet much higher, modelled erosion
    potential (Fig.~\ref{fig:powerlaws}b). The erosion law based on measured
    suspended sediment load during a surge of Variegated Glacier
    \citep[${\dot{e}=1\times10^{-4}u_\mathrm{b}^{1}}$,][]{Humphrey.Raymond.1994}
    results in similarly high values, but a less localized pattern of cumulative
    erosion potential (Fig.~\ref{fig:powerlaws}c). Finally, The erosion law
    based on a global compilation of glacier velocities and erosion rates
    \citep[${\dot{e}=1.665\times10^{-1}u_\mathrm{b}^{0.6459}}$,][]{Cook.etal.2020}
    yields an even flatter pattern of cumulative erosion potential
    (Fig.~\ref{fig:powerlaws}d).

    It may seem surprising that the erosion law calibrated on Patagonian and
    Antarctic Peninsula tidewater glaciers \citep{Koppes.etal.2015} yields a
    more realistic glacial cycle erosion potential in the Alps than those based
    on terrestrial
    glaciers \citep{Humphrey.Raymond.1994, Herman.etal.2015, Cook.etal.2020},
    themselves resulting in kilometre-scale integrated erosion even
    higher than the Alpine relief itself. However, during the Last Glacial
    Maximum and much of the last glacial cycle, Alpine paleoglaciers were
    closer in size, slope, and climatic context to the present-day glaciers of
    Patagonia and the Antarctic Peninsula \citep{Koppes.etal.2015} than to
    Franz-Joseph Glacier \citep{Herman.etal.2015} and many of the glaciers
    included in the global compilation by \citet{Cook.etal.2020}. Indeed,
    atmospheric
    temperatures have been found to be a strong driver of erosion in studies
    involving several glaciers \citep{Koppes.etal.2015, Cook.etal.2020}.

    That said, the magnitude of the modelled erosion rates can hardly be used
    as a criteria to rule out this or that erosion law for the glacial Alps, as
    all these laws are affected with very large uncertainties. Thus it can not
    be excluded that a linear or sub-linear erosion-law
    \citep[e.g.,][]{Cook.etal.2020} would lead to a less localized pattern of
    erosion (Fig.~\ref{fig:powerlaws}c and~d), and a stronger correlation
    between modelled ice volume and erosion (Fig.~\ref{fig:powerlaws}k and~l).
    Even then, the intrinsically large variations in modelled glacier sliding
    velocities yield cumulative glacial-cycle erosion two to three orders of
    magnitude higher in the Alpine valley than near the mountain tops
    (Fig.~\ref{fig:powerlaws}d).

    \todo{Insert discussion of sediments here.}
